---
title: "Introduction to bivariate polar plots using `polarplotr`"
author: "**David Carslaw**"
date: "`r format(Sys.Date(), '%e %B %Y' )`"
output: 
  github_document: 
    toc: TRUE
vignette: >
  %\VignetteIndexEntry{polarplotr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r echo=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  eval = FALSE,
  cache = TRUE,
  dpi=300
)
```

## Introduction

This document considers the use of bivariate polar plots applied to air pollution problems. Since the development of the **openair** package, which first made available the `polarPlot` function, there have been many further developments. These include:

1. Clustering of polar plot concetration surfaces (Carslaw and Beevers, 2013).
1. Development of a bivarate Conditional Probability Function (Uria-Tellaetxe and Carslaw, 2014).
1. Extension to considering two pollutant species and a range of pair-wise statistics (Grange et al., 2016).

This document summarises the main use of the polar plot functions with examples using real air quality data,

The `polarPlot` function plots a *bivariate polar plot* of
concentrations. Concentrations are shown to vary by wind speed and
wind direction. 

For many, maybe most situations, increasing wind speed generally
results in lower concentrations due to increased dilution through
advection and increased mechanical turbulence. There are, however,
many processes that can lead to interesting concentration-wind speed
dependencies and we will provide a more theoretical treatment of this
in due course. However, below are a few reasons why concentrations can
change with increasing wind speeds.


- Buoyant plumes from tall stacks can be brought down to
  ground-level resulting in high concentrations under high wind speed
  conditions.

- Particle suspension increases with increasing wind speeds
  e.g. \pmten from spoil heaps and the like.

- 'Particle' suspension can be important close to coastal areas
  where higher wind speeds generate more sea spray.

- The wind speed dependence of concentrations in a street canyon
  can be very complex: higher wind speeds do not always results in
  lower concentrations due to re-circulation. Bivariate polar plots are
  very good at revealing these complexities.

- As Carslaw et al. (2006) showed, aircraft emissions have an unusual
  wind speed dependence and this can help distinguish them from other
  sources. If several measurement sites are available, polar plots can
  be used to triangulate different sources.

- Concentrations of NO~2~ can increase with increasing wind speed
  --- or at least not decline steeply due to increased mixing. This
  mixing can result in O~3~ -rich air converting NO to NO~2~.



## Example of use

For these examples use will be made of air quality data from the UK and surface meteorological data from the `worldmet` package, available [here](https://github.com/davidcarslaw/worldmet). 

```{r eval=TRUE}
library(openair)
library(polarplotr)
library(dplyr)
library(worldmet)

# import data from London North Kensington
north_ken <- importAURN(site = "kc1", year = 2013)

# import met data for site Heathrow London (which is the default)
met <- importNOAA(year = 2013)

# merge together, drop model met from north_ken
north_ken <- inner_join(select(north_ken, -ws, -wd),
                        met, by = "date")
```

We will also use data from a kerbside site (Marylebone Road).

```{r eval=TRUE}
# import data from London Marylebone Road
mary <- importAURN(site = "my1", year = 2013)

# merge together, drop model met from north_ken
mary <- inner_join(select(mary, -ws, -wd),
                        met, by = "date")
```

Basic polar plot.

```{r eval = TRUE,fig.width=5,fig.height=4.5,out.width=300}
polarPlot(north_ken, pollutant = "nox", col = "jet")
```

Useful options for `polarPlot` include:

- `x`, the variable for the radial axis (wind speed by default). The radial axis can be any numeric variable. The choice of variable depends on teh application but is usually chosen to help differentiate sources in some way e.g. air temperature or a measure of atmospheric stability.
- `limits` to set the colour scale limits e.g. `limits = c(0, 100)`. This is useful for ensuring the same scale across different plots.
- `k`, which controls the smoothing parameter. Generally it is not necessary to adjust `k` but if there are too few data it can be difficult to fit a smooth surface. Reducing k e.g. to 50 or below might help produce a plot.

## Conditional Probability Function

## Comparing two pollutants

## References

1. Carslaw, D.C. and K. Ropkins, (2012). openair --- an R package for air quality data analysis. Environmental Modelling & Software. Volume 27-28, 52-61.

1.  Carslaw, D.C., Beevers, S.D, Ropkins, K and M.C. Bell (2006). 
 Detecting and quantifying aircraft and other on-airport 
 contributions to ambient nitrogen oxides in the vicinity of a 
 large international airport.  Atmospheric Environment. 40/28 pp 
 5424-5434.
 
1.  Carslaw, D.C., & Beevers, S.D. (2013). Characterising and 
 understanding emission sources using bivariate polar plots and 
 k-means clustering. Environmental Modelling & Software, 40, 
 325-329. doi:10.1016/j.envsoft.2012.09.005
 
1.  Uria-Tellaetxe, I. and D.C. Carslaw (2014). Source identification 
 using a conditional bivariate Probability function. Environmental Modelling & Software, Vol. 59, 1-9. [open access version](http://www.sciencedirect.com/science/article/pii/S1364815214001339).
 
1.  Westmoreland, E.J., N. Carslaw, D.C. Carslaw, A. Gillah and E.
 Bates (2007).  Analysis of air quality within a street canyon 
 using statistical and dispersion modelling techniques. Atmospheric
 Environment. Vol.  41(39), pp. 9195-9205.
 
1. Grange, S. K., Lewis, A. C. and D. C. Carslaw (2016). Source apportionment 
 advances with bivariate polar plots, correlation, and regression techniques.
 Atmospheric Environment. 145, 128-134. [open access version](http://www.sciencedirect.com/science/article/pii/S1352231016307166).