---
title: "**polarplotr**"
output: 
  html_document: 
    theme: cosmo
    fig_height: 4
    fig_width: 12
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Infomation

**polarplotr** is an R package which allows to plot bivariate polar plots for air quality time-series data. **polarplotr** started life in the [**openair**](https://github.com/davidcarslaw/openair) R package, but has been pulled from this original package and is now developed as its own package. **polarplotr**'s main function is `polarPlot` and is very useful when exploring and analysis pollutant sources and patterns. 

## Citation

Please cite the **polarplotr** package like this:

```
David Carslaw and Stuart Grange (2016). polarplotr: Functions to plot polar-plots. https://github.com/skgrange/polarplotr
```

For for LaTeX users:

```
@Manual{,
  title = {polarplotr: Functions to plot polar-plots},
  author = {David Carslaw and Stuart Grange},
  year = {2016},
  url = {https://github.com/skgrange/polarplotr},
}
```

## Some examples

A simple example using York Fishergate, a roadside site in York, North Yorkshire, United Kingdom. 

```{r, echo = TRUE, cache = TRUE, message = FALSE, fig.width = 7}
# Load packages
library(openair)
library(polarplotr)

# Get some data
data_yk11 <- importAURN("yk11", 2014)

# Plot
polarPlot(data_yk11, "nox")
polarPlot(data_yk11, "pm10")
```
